---
title: "Sampling"
author: "Oluchi Nnadi"
format: html
execute:
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(primer.data)
```


This bar shows a 95% confidence interval for the mean age of the shaming tibble based on a sample of 50 observations. The dashed red line represents the true mean age of the entire dataset. The fact that the confidence interval includes the true mean suggests that our sample provides a reasonable estimate of the population parameter.
```{r}
set.seed(10)
truth <- mean(shaming$age, na.rm = TRUE)

my_sample <- sample_n(shaming, 50)
sample_mean <- mean(my_sample$age, na.rm = TRUE)
sample_se <- sd(my_sample$age, na.rm = TRUE) / sqrt(nrow(my_sample))

print("The true mean age of the shaming tibble:")
print(truth)
print("The  mean age of our sample (n = 50):")
print(sample_mean)
print("The standard error of our sample:")
print(sample_se)

tibble(
  estimate = sample_mean,
  lower = sample_mean - 1.96 * sample_se,
  upper = sample_mean + 1.96 * sample_se
) |>
  ggplot(aes(x = estimate, y = 0)) +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = lower, xmax = upper), orientation = "y", width = 0.1) +
  geom_vline(aes(xintercept = truth), color = "red", linetype = "dashed") +
  labs(
    title = "Sample mean estimate with a 95% error bar",
    subtitle = "The sample mean is close to the population truth, suggesting this sample provides a reasonable estimate.",
    x = "Estimate",
    y = ""
  ) +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
```


```{r}
#| cache: true
set.seed(10)

samples <- tibble(replicate = 1:1000) |>
  mutate(
    sample = map(replicate, ~ sample_n(shaming, 50)),
    sample_mean = map_dbl(sample, ~ mean(.x$age, na.rm = TRUE)),
    sample_se = map_dbl(sample, ~ sd(.x$age, na.rm = TRUE) / sqrt(nrow(.x))),
    lower = sample_mean - 1.96 * sample_se,
    upper = sample_mean + 1.96 * sample_se
  ) |>
  select(replicate, sample_mean, sample_se, lower, upper)
```


This plot displays the first 100 sample means with their corresponding 95% confidence intervals. The dashed red line represents the true mean age of the entire dataset. Most of the intervals contain the true mean, demonstrating that repeated random samples often recover the population mean.
```{r}
set.seed(10)

samples <- tibble(replicate = 1:1000) |>
  mutate(
    sample = map(replicate, ~ sample_n(shaming, 50)),
    sample_mean = map_dbl(sample, ~ mean(.x$age, na.rm = TRUE)),
    sample_se = map_dbl(sample, ~ sd(.x$age, na.rm = TRUE) / sqrt(nrow(.x))),
    lower = sample_mean - 1.96 * sample_se,
    upper = sample_mean + 1.96 * sample_se
  ) |>
  select(replicate, sample_mean, sample_se, lower, upper)

samples |>
  slice_head(n = 100) |>
  ggplot(aes(y = factor(replicate))) +
  geom_point(aes(x = sample_mean), size = 2) +
  geom_errorbar(aes(xmin = lower, xmax = upper), orientation = "y", width = 0.1) +
  geom_vline(aes(xintercept = truth), color = "red", linetype = "dashed") +
  labs(
    title = "First 100 sample means with 95% confidence intervals",
    subtitle = "Most intervals contain the truth, showing that repeated random samples often recover the population mean.",
    x = "Estimate",
    y = "Sample"
  ) +
  scale_y_discrete(breaks = seq(10, 100, by = 10)) +
  theme_minimal()
```

This is the fraction of the 1000 samples that contain the true mean age of the shaming tibble. A coverage fraction close to 0.95 indicates that our confidence intervals are well-calibrated and provide a reliable estimate of the population parameter.
```{r}
coverage_fraction <- mean(samples$lower < truth & samples$upper > truth)
print(coverage_fraction)
```


These histograms show the age distributions of a biased sample and an unbiased sample, each containing 50 observations. The biased sample is skewed toward older voters, while the unbiased sample more accurately reflects the overall population distribution. This illustrates how sampling methods can influence the representativeness of a sample.
```{r}
set.seed(10)

weights <- if_else(shaming$age > 40, 3, 1)
selected_rows <- sample.int(nrow(shaming), size = 50, replace = TRUE, prob = weights)
biased_sample <- shaming[selected_rows, ]

set.seed(11)
unbiased_sample <- sample_n(shaming, 50)

bind_rows(
  biased_sample |> mutate(sample_type = "biased"),
  unbiased_sample |> mutate(sample_type = "unbiased")
) |>
  ggplot(aes(x = age, fill = sample_type)) +
  geom_histogram(bins = 15, alpha = 0.6, position = "identity") +
  facet_wrap(~ sample_type, ncol = 2) +
  labs(
    title = "Age distributions of biased and unbiased samples",
    subtitle = "The biased sample is visibly skewed toward older voters",
    x = "Age",
    y = "Count"
  ) +
  theme_minimal()

biased_sample_mean <- mean(biased_sample$age, na.rm = TRUE)
biased_sample_se <- sd(biased_sample$age, na.rm = TRUE) / sqrt(nrow(biased_sample))

print("The mean age of the biased sample (it's 5 years above the true mean):")
print(biased_sample_mean)
print("The standard error of the biased sample:")
print(biased_sample_se)
```

```{r}
set.seed(10)

biased_samples <- tibble(replicate = 1:1000) |>
  mutate(
    sample = map(replicate, ~ {
      weights <- if_else(shaming$age > 40, 3, 1)
      selected_rows <- sample.int(nrow(shaming), size = 50, replace = TRUE, prob = weights)
      shaming[selected_rows, ]
    }),
    sample_mean = map_dbl(sample, ~ mean(.x$age, na.rm = TRUE)),
    sample_se = map_dbl(sample, ~ sd(.x$age, na.rm = TRUE) / sqrt(nrow(.x))),
    lower = sample_mean - 1.96 * sample_se,
    upper = sample_mean + 1.96 * sample_se
  ) |>
  select(replicate, sample_mean, sample_se, lower, upper)
```

This plot displays the first 100 biased sample means with their corresponding 95% confidence intervals. The dashed red line represents the true mean age of the entire dataset. Many of the intervals do not contain the true mean, indicating that the biased sampling scheme skews the estimates toward older voters.
```{r}
set.seed(10)

biased_samples <- tibble(replicate = 1:1000) |>
  mutate(
    sample = map(replicate, ~ {
      weights <- if_else(shaming$age > 40, 3, 1)
      selected_rows <- sample.int(nrow(shaming), size = 50, replace = TRUE, prob = weights)
      shaming[selected_rows, ]
    }),
    sample_mean = map_dbl(sample, ~ mean(.x$age, na.rm = TRUE)),
    sample_se = map_dbl(sample, ~ sd(.x$age, na.rm = TRUE) / sqrt(nrow(.x))),
    lower = sample_mean - 1.96 * sample_se,
    upper = sample_mean + 1.96 * sample_se
  ) |>
  select(replicate, sample_mean, sample_se, lower, upper)

biased_samples |>
  slice_head(n = 100) |>
  ggplot(aes(y = factor(replicate))) +
  geom_point(aes(x = sample_mean), size = 2) +
  geom_errorbar(aes(xmin = lower, xmax = upper), orientation = "y", width = 0.1) +
  geom_vline(aes(xintercept = truth), color = "red", linetype = "dashed") +
  labs(
    title = "First 100 biased sample means with 95% confidence intervals",
    subtitle = "The biased sampling scheme pulls many intervals toward older voters",
    x = "Estimate",
    y = "Sample"
  ) +
  theme_minimal()
```

